---
title: "P2_Buffer_intersect"
author: "Bruna Amaral"
date: "2022-12-09"
output: html_document
---

Code to get the area of each park, used to create a buffer around it, and crop land use rasters with the buffer (and change its resolution for landscapemetrics analysis - x and y resolution must be the same)

 Input:  - park_raster/nps_boundary/nps_boundary.shp: shape file with the boundaries of US parks
         - park_raster/{park}_{year}_lc.rds
 Output: - park_raster/{park}_pb.rds


```{r, message=FALSE}
## load packages ---------------------------------------------------------------------
library(dplyr)
library(raster)
library(sp)
library(rgdal)
library(rgeos)
library(maptools)
library(FedData)
library(data.table)
library(tidyverse)
library(stars)
library(sf)
library(glue)
library(terra)

setwd("~/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/MSU_courses/IBIO_SpaEco/Project")
```

Get the area of each park in a file
```{r, eval=FALSE}
## get park area maps -----------------------------------------------------------------
park_bound <- readOGR(file.path("park_raster/nps_boundary/nps_boundary.shp")) #st_read("park_raster/nps_boundary/nps_boundary.shp")

# ACAD
#acad_2019_lc <- read_rds("park_raster/acad_2019_lc.rds")  # plot shape on raster to see if things are lining up
pb_acad <- subset(park_bound, UNIT_CODE == "ACAD")
write_rds(pb_acad, file = "park_raster/acad_pb.rds")
# plot(acad_2019_lc)
# plot(pb_acad, add = T, main = "ACAD")

# ELRO
#elro_2019_lc <- read_rds("park_raster/elro_2019_lc.rds")
pb_elro <- subset(park_bound, UNIT_CODE == "ELRO")
write_rds(pb_elro, file = "park_raster/elro_pb.rds")
# plot(elro_2019_lc)
# plot(pb_elro, add = T, main = "ELRO")

# HOFR
#hofr_2019_lc <- read_rds("park_raster/hofr_2019_lc.rds")
pb_hofr <- subset(park_bound, UNIT_CODE == "HOFR")
write_rds(pb_hofr, file = "park_raster/hofr_pb.rds")
# plot(hofr_2019_lc)
# plot(pb_hofr, add = T, main = "HOFR")

# MABI
#mabi_2019_lc <- read_rds("park_raster/mabi_2019_lc.rds")
pb_mabi <- subset(park_bound, UNIT_CODE == "MABI")
write_rds(pb_mabi, file = "park_raster/mabi_pb.rds")
# plot(mabi_2019_lc)
# plot(pb_mabi, add = T, main = "MABI")

# MIMA
#mima_2019_lc <- read_rds("park_raster/mima_2019_lc.rds")
pb_mima <- subset(park_bound, UNIT_CODE == "MIMA")
write_rds(pb_mima, file = "park_raster/mima_pb.rds")
# plot(mima_2019_lc)
# plot(pb_mima, add = T, main = "MIMA")

# MORR
#morr_2019_lc <- read_rds("park_raster/morr_2019_lc.rds")
pb_morr <- subset(park_bound, UNIT_CODE == "MORR")
write_rds(pb_morr, file = "park_raster/morr_pb.rds")
# plot(morr_2019_lc)
# plot(pb_morr, add = T, main = "MORR")

# SAGA
#saga_2019_lc <- read_rds("park_raster/saga_2019_lc.rds")
pb_saga <- subset(park_bound, UNIT_CODE == "SAGA")
write_rds(pb_saga, file = "park_raster/saga_pb.rds")
# plot(saga_2019_lc, xlim = c(-72.5, -72.2), ylim = c(43.44, 43.6))
# plot(pb_saga, add = T, main = "SAGA")

# SARA
#sara_2019_lc <- read_rds("park_raster/sara_2019_lc.rds")
pb_sara <- subset(park_bound, UNIT_CODE == "SARA")
write_rds(pb_sara, file = "park_raster/sara_pb.rds")
# plot(sara_2019_lc)
# plot(pb_sara, add = T, main = "SARA")

# SAIR
#sair_2019_lc <- read_rds("park_raster/sair_2019_lc.rds")
pb_sair <- subset(park_bound, UNIT_CODE == "SAIR")
write_rds(pb_sair, file = "park_raster/sair_pb.rds")
# plot(sair_2019_lc)
# plot(pb_sair, add = T, main = "SAIR")

# VAMA
#vama_2019_lc <- read_rds("park_raster/vama_2019_lc.rds")
pb_vama <- subset(park_bound, UNIT_CODE == "VAMA")
write_rds(pb_vama, file = "park_raster/vama_pb.rds")
# plot(vama_2019_lc)
# plot(pb_vama, add = T, main = "VAMA")

# WEFA
#wefa_2019_lc <- read_rds("park_raster/wefa_2019_lc.rds")
pb_wefa <- subset(park_bound, UNIT_CODE == "WEFA")
write_rds(pb_wefa, file = "park_raster/wefa_pb.rds")
# plot(wefa_2019_lc)
# plot(pb_wefa, add = T, main = "WEFA")
```

Create a buffer around every park to crop the land use map accordingly
```{r, eval = FALSE}
sf_use_s2(FALSE)

# ACAD
geom_acad <- st_as_sfc(pb_acad)
buf_acad <- st_buffer(geom_acad, dist = 0.1)
write_rds(buf_acad, file = "park_raster/buf_acad.rds")
#plot(acad_2019_lc)
#, xlim = c(-68.7,-68), ylim = c(44.2,44.5))
# plot(geom_acad, add = TRUE)
# plot(buf_acad, border = 'red', add = T)

# ELRO
geom_elro <- st_as_sfc(pb_elro)
buf_elro <- st_buffer(geom_elro, dist = 0.1)
write_rds(buf_elro, file = "park_raster/buf_elro.rds")
# plot(elro_2019_lc)
# plot(geom_elro, add = TRUE)
# plot(buf_elro, border = 'red', add = T)

# HOFR
geom_hofr <- st_as_sfc(pb_hofr)
buf_hofr <- st_buffer(geom_hofr, dist = 0.1)
write_rds(buf_hofr, file = "park_raster/buf_hofr.rds")
# plot(hofr_2019_lc)
# plot(geom_hofr, add = TRUE)
# plot(buf_hofr, border = 'red', add = T)

# MABI
geom_mabi <- st_as_sfc(pb_mabi)
buf_mabi <- st_buffer(geom_mabi, dist = 0.1)
write_rds(buf_mabi, file = "park_raster/buf_mabi.rds")
# plot(mabi_2019_lc)
# plot(geom_mabi, add = TRUE)
# plot(buf_mabi, border = 'red', add = T)

# MIMA
geom_mima <- st_as_sfc(pb_mima)
buf_mima <- st_buffer(geom_mima, dist = 0.1)
write_rds(buf_mima, file = "park_raster/buf_mima.rds")
# plot(mima_2019_lc)
# plot(geom_mima, add = TRUE)
# plot(buf_mima, border = 'red', add = T)

# MORR
geom_morr <- st_as_sfc(pb_morr)
buf_morr <- st_buffer(geom_morr, dist = 0.1)
write_rds(buf_morr, file = "park_raster/buf_morr.rds")
# plot(morr_2019_lc)
# plot(geom_morr, add = TRUE)
# plot(buf_morr, border = 'red', add = T)

# SAGA
geom_saga <- st_as_sfc(pb_saga)
buf_saga <- st_buffer(geom_saga, dist = 0.1)
write_rds(buf_saga, file = "park_raster/buf_saga.rds")
# plot(saga_2019_lc)
# plot(geom_saga, add = TRUE)
# plot(buf_saga, border = 'red', add = T)

# SAIR
geom_sair <- st_as_sfc(pb_sair)
buf_sair <- st_buffer(geom_sair, dist = 0.1)
write_rds(buf_sair, file = "park_raster/buf_sair.rds")
# plot(sair_2019_lc) #, xlim = c(-71.02,-71), ylim = c(42.46, 42.48))
# plot(geom_sair, add = TRUE)
# plot(buf_sair, border = 'red', add = T)

# SARA
geom_sara <- st_as_sfc(pb_sara)
buf_sara <- st_buffer(geom_sara, dist = 0.1)
write_rds(buf_sara, file = "park_raster/buf_sara.rds")
# plot(sara_2019_lc)
# plot(geom_sara, add = TRUE)
# plot(buf_sara, border = 'red', add = T)

# VAMA
geom_vama <- st_as_sfc(pb_vama)
buf_vama <- st_buffer(geom_vama, dist = 0.1)
write_rds(buf_vama, file = "park_raster/buf_vama.rds")
# plot(vama_2019_lc)
# plot(geom_vama, add = TRUE)
# plot(buf_vama, border = 'red', add = T)

# WEFA
geom_wefa <- st_as_sfc(pb_wefa)
buf_wefa <- st_buffer(geom_wefa, dist = 0.1)
write_rds(buf_wefa, file = "park_raster/buf_wefa.rds")
# plot(wefa_2019_lc)
# plot(geom_wefa, add = TRUE)
# plot(buf_wefa, border = 'red', add = T)
```

Crop the different year and park maps
```{r, eval=FALSE}
pn <- read_rds("park_raster/key_park.rds")

park_list <- pn$parks %>% str_to_lower()
#park_list <- park_list[-9]

years <- c(2001, 2006, 2008, 2011, 2013, 2016, 2019)

lcc.proj <- CRS("+proj=lcc +lat_0=41.75 +lon_0=-120.5 +lat_1=43 +lat_2=45.5 +x_0=400000 +y_0=0 +datum=NAD83 +units=ft +no_defs")

# function to crop different park and years
buf_cut <- function(park, year) {
  
  buf <- readRDS(file = glue("park_raster/buf_{park}.rds"))
  if(park == "hofr" & year == 2011) {
    # for some reason hofr_2011_lc does not work on loop
    hofr_2011_lc <- raster("park_raster/HOFR/NLCD_2011_Land_Cover_L48_20210604_FXrB2kVAzgyCFr4fLALu.tiff")
    hofr_2011_lc <- projectRaster(hofr_2011_lc,
                                  crs = CRS("+proj=longlat +datum=NAD83 +no_defs"))
    park_map_year <- hofr_2011_lc
  } else if (park == "mabi" & year == 2016){
    # for some reason mabi_2016_lc does not work on loop
    mabi_2016_lc <- raster("park_raster/MABI/NLCD_2016_Land_Cover_L48_20210604_NQqfiy9JRFKBGlEHX0pL.tiff")
    mabi_2016_lc <- projectRaster(mabi_2016_lc,
                                  crs = CRS("+proj=longlat +datum=NAD83 +no_defs"))
    park_map_year <- mabi_2016_lc
  } else { 
    park_map_year <- readRDS(file = glue("park_raster/{park}_{year}_lc.rds"))
  }
  
  buf_ext <- raster::extent(st_as_sf(buf))
  int_c <- crop(park_map_year, buf_ext) 
  int <- mask(int_c, st_as_sf(buf)) 
  
  saveRDS(int, file = glue("park_raster/{park}{year}_rasbuf.rds"))
  
  int2 <- int
  raster::values(int2) <- raster::values(int2) %>% round(digits = 0)
  # selecting only forest
  int2[int2 < 41] <- 0 # make everything below 41 zero
  int2[int2 > 43] <- 0 # make everything above 43 zero
  int2[int2 == 41 | int2 == 42 | int2 == 43] <- 1
  
  # resolution on x and y is different
  # raster::res(hood_for)
  int3 <- projectRaster(int2, crs = lcc.proj, res=c(50,50))
  int3[int3 > 0.5] <- 1
  int3[int3 < 0.5] <- 0
  saveRDS(int3, file = glue("park_raster/{park}{year}_eqres.rds"))

}

for(i in 1:length(park_list)) {
  for(j in 1:length(years)){
    park <- park_list[i]
    year <- years[j]
    buf_cut(park, year) 
  }
}

```

Example of the steps of getting the park area, creating the buffer and croping the landscape layer 
```{r, echo=TRUE}
park <- "acad"
year <- 2019

raster::plot(readRDS(glue("~/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/MSU_courses/IBIO_SpaEco/Project/park_raster/{park}_pb.rds")))
raster::plot(readRDS(glue("~/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/MSU_courses/IBIO_SpaEco/Project/park_raster/buf_{park}.rds")))
raster::plot(readRDS(glue("~/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/MSU_courses/IBIO_SpaEco/Project/park_raster/{park}{year}_eqres.rds")))


```