---
title: "P3"
author: "Bruna Amaral"
date: "2022-12-09"
output: html_document
---

Script to get the park buffer areas with forest class and calculate landscape metrics

Input: 

park_raster/key_park.rds - list of park names  

park_raster/{park}{year}_eqres.rds - raster file with forest cover classification
       
park_raster/{park}{year}_rasbuf.rds - raster file with all landscape cover classes
       
Output: 

park_raster/{park}{year}_metricsRAW.rds - all landscape metrics
       
park_raster/{park}{year}_metrics.rds - selected landscape metrics for parks
       
metrics_tab.rds - matrix for all parks with landscape metrics
       
metrics_tab2.rds - matrix for all parks with some extra landscape metrics


```{r, message=FALSE}
## load packages ------------------------------------------------------------
library(dplyr)
library(ggplot2)
library(raster)
library(sp)
library(rgdal)
library(rgeos)
library(maptools)
library(landscapemetrics)
library(landscapetools)
library(FedData)
#library(GISTools) # dont work - ok
library(data.table)
library("landscapemetrics")
library("landscapetools")
library("remotes")
library(prettymapr)
library(cowplot)
library(dismo)
library(tidyverse)
library(glue)

```

```{r, eval= F}
# park list
pn <- read_rds("park_raster/key_park.rds")

park_list <- pn$parks %>% str_to_lower()

years <- c(2001, 2006, 2008, 2011, 2013, 2016, 2019)

metrics_tab <- tidyr::crossing(park_list, years) %>% 
  mutate(IJI = as.numeric(NA),
         CI = as.numeric(NA),
         TCA = as.numeric(NA),
         TE = as.numeric(NA),
         NP = as.numeric(NA))

patch_design <- function(park, year) {
  
  hood_for2 <- read_rds(glue("park_raster/{park}{year}_eqres.rds"))
  
  hood_metrics <- calculate_lsm(hood_for2, what = "patch")
  
  saveRDS(hood_metrics, file = glue("park_raster/{park}{year}_metricsRAW.rds"))
  
   hood_for_metrics_hist <- hood_metrics_for %>%
     dplyr::select(value, metric) %>%
     na.omit()
   saveRDS(hood_for_metrics_hist, file = glue("park_raster/{park}{year}_metrics.rds"))
  
}

for(i in 1:length(park_list)) {
  for(j in 1:length(years)){
    park <- park_list[i]
    year <- years[j]
    
    # basic metrics
    patch_design(park, year) 
    
    # extra landscape metrics:
    
    hood_for2 <- read_rds(glue("park_raster/{park}{year}_eqres.rds"))
    hood_for <- read_rds(glue("park_raster/{park}{year}_rasbuf.rds"))
    
    row_numb <- which(metrics_tab$park_list == park & metrics_tab$years == year, arr.ind = T)
    
    # Interspersion and Juxtaposition index (percent)
    hood_metrics_iji <- lsm_l_iji(hood_for)
    metrics_tab$IJI[row_numb] <- hood_metrics_iji$value
    
    # Contagion index (percent)
    hood_metrics_con <- lsm_l_contag(hood_for2)
    metrics_tab$CI[row_numb] <- hood_metrics_con$value
    
    # Total core area (ha)
    hood_metrics_tca <- lsm_l_tca(hood_for2)
    metrics_tab$TCA[row_numb] <- hood_metrics_tca$value
    
    # Total edge (m)
    hood_metrics_te <- lsm_c_te(hood_for2)
    metrics_tab$TE[row_numb] <- hood_metrics_te %>% filter(class == 1) %>% dplyr::select(value) %>% pull()
    
    # number of patches
    hood_metrics_np <- lsm_l_np(hood_for2)
    metrics_tab$NP[row_numb] <- hood_metrics_np$value
    
    saveRDS(metrics_tab, file = "metrics_tab1.rds")
  }
}
```


Get some extra landscape metrics
```{r, eval= F}
# even more patch metrics:
metrics_tab <- tidyr::crossing(park_list, years) %>% 
  mutate(cai = as.numeric(NA),
         coh = as.numeric(NA),
         enn = as.numeric(NA),
         cai = as.numeric(NA),
         nlsi = as.numeric(NA),
         clu = as.numeric(NA))

for(i in 1:length(park_list)) {
  for(j in 1:length(years)){
    park <- park_list[i]
    year <- years[j]
    
    #patch_design(park, year) 
    
    # extra landscape metrics:
    hood_for2 <- read_rds(glue("park_raster/{park}{year}_eqres.rds"))
    hood_for <- read_rds(glue("park_raster/{park}{year}_rasbuf.rds"))
    
    row_numb <- which(metrics_tab$park_list == park & metrics_tab$years == year, arr.ind = T)
    
    # CAI_MN - Mean Core Area Index (CORE AREA)
    hood_metrics_cai <- lsm_l_cai_mn(hood_for2)
    metrics_tab$cai[row_numb] <- hood_metrics_cai$value
    
    # CONNECT - Connectance Index - not available in package (CONNECTIVITY) - not available
    # COHESION - Patch Cohesion Index
    hood_metrics_coh <- lsm_l_cohesion(hood_for2)
    metrics_tab$coh[row_numb] <- hood_metrics_coh$value
    
    # ECON_AM - Area Weighted Mean Edge Contrast Index - not available (CONTRAST) - not available
    # TECI - Total Edge Contrast Index (CONTRAST) - not available
    
    # PROX_CV - Coefficient of Variation of Proximity Index (PROXIMITY/ISOLATION) - not available
    # ENN_MN - Mean Euclidian Nearest Neighbour Index
    hood_metrics_enn <- lsm_l_enn_mn(hood_for2)
    metrics_tab$enn[row_numb] <- hood_metrics_enn$value
    
    # AI - Aggregation Index (CONTAGION/INTERSPERSION)
    hood_metrics_ai <- lsm_l_ai(hood_for2)
    metrics_tab$ai[row_numb] <- hood_metrics_ai$value
    
    # nLSI - Normalized Landscape Shape Index (AREA/EDGE/DENSITY)
    hood_metrics_nlsi <- lsm_c_nlsi(hood_for2)
    metrics_tab$nlsi[row_numb] <- hood_metrics_nlsi$value
    
    # PAFRAC - Perimeter Area Fractal Dimension (SHAPE)
    hood_metrics_pafrac <- lsm_l_pafrac(hood_for2)
    metrics_tab$pafrac[row_numb] <- hood_metrics_pafrac$value
    
    # CLUMPY - Clumpy Index (CONTAGION/INTERSPERSION)
    hood_metrics_clu <- lsm_c_clumpy(hood_for2)
    metrics_tab$clu[row_numb] <- hood_metrics_clu$value
    
    saveRDS(metrics_tab, file = "metrics_tab2.rds")
  }
}
```

Join two metrics tabs
```{r, eval=FALSE}
metrics_tab1 <- readRDS("metrics_tab1.rds")
metrics_tab2 <- readRDS("metrics_tab2.rds")
metrics_tab <- left_join(metrics_tab1, metrics_tab2, by = c(park_list, years))

saveRDS(metrics_tab, file = "metrics_tab.rds")


```

Check the metrics table that was created
```{r}
metrics_tab <- readRDS("metrics_tab.rds")
print(metrics_tab)

```
