---
title: "P4_metrics_plots"
author: "Bruna Amaral"
date: "2022-12-09"
output: html_document
---

Script to create plots of different landscape metrics varying by park and by year

Input: 

park_raster/key_park.rds - table with park names

park_raster/{park}{year}_metrics.rds - matrix with first round of landscape metrics for parks

metrics_tab.rds - matrix with extra landscape metrics calculated for the parks

Output: 

park_raster/{park}_mets.rds - matrix with organized landscape metrics for each park

plots for each metric (park and year variation)

```{r, message = F}
# load packages ---------------------------------------------------------------------
library(lattice)
library(tidyverse)
library(glue)

knitr::opts_chunk$set(fig.width=25*1.3, fig.height=4*1.3)

```

```{r}
# park list
pn <- read_rds("park_raster/key_park.rds")

park_list <- pn$parks %>% str_to_lower()

years <- c(2001, 2006, 2008, 2011, 2013, 2016, 2019)

## create matrix with metrix for all years and parks

pat_met <- expand.grid(park_list, years)
colnames(pat_met) <- c("park", "year")
pat_met <- pat_met %>% 
  as_tibble() %>% 
  arrange(park) %>% 
  mutate("metric" = NA,
         "value" = NA)

for(i in 1:length(park_list)) {
  park <- park_list[i]
  for(j in 1:length(years)){  
    year <- years[j]
    met <- read_rds(glue("park_raster/{park}{year}_metrics.rds"))
    met <- met %>% 
      mutate(park = park,
             year =year) %>% 
    relocate(park, year, metric, value)
    
    if(j == 1){met_p <- met
    } else {
      met_p <- rbind(met_p, met)
    }
    assign(glue("{park}_mets"), met_p)
    write_rds(met_p, glue("park_raster/{park}_mets"))
  }
}

mets_all <- rbind(acad_mets, elro_mets, hofr_mets,
                  mabi_mets, mima_mets, morr_mets,
                  saga_mets, sair_mets, sara_mets,
                  vama_mets, wefa_mets) %>% 
  mutate(park = as.factor(park)) %>% 
  filter(year > 2005)

unique(mets_all$metric)

metrics_tab <- read_rds("metrics_tab.rds")

```


## Area
```{r}
# area -------------------------
area_p <- mets_all %>% filter(metric == "area") %>% 
  group_by(park, year) %>% 
  summarise(total_area = sum(value)) %>% 
  mutate(total_area_z = scale(total_area, center = FALSE, scale = TRUE),
         year_z = scale(year, center = TRUE, scale = TRUE))

sd(area_p$total_area)/1000
mean(area_p$total_area)

#svg(glue("area_z.#svg"), 
#    width = 25, height = 4)
ggplot(area_p,# %>% filter(park == "acad"),
       aes(y = total_area_z, x = year, group= park, col=park)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park, scales = "free") +
  theme(text = element_text(size = 25), legend.position = "none")
#dev.off()

#svg(glue("area.#svg"), 
#    width = 25, height = 4)
ggplot(area_p,# %>% filter(park == "acad"),
       aes(y = total_area, x = year, group= park, col=park)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park, scales = "free")+
  theme(text = element_text(size = 25),legend.position = "none")

#dev.off()
```


## Number of core areas
```{r}
# Number of core areas -------------------------------------------------------
ncore_p2 <- mets_all %>% filter(metric == "ncore") %>% 
  group_by(park, year) %>% 
  tally() %>% 
  mutate(n_z = scale(n, center = T, scale = T))

#svg(glue("ncore_z.#svg"), 
#    width = 25, height = 4)
ggplot(ncore_p2,aes(y = n_z, x = year, group= park, col=park)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

#svg(glue("ncore.#svg"), 
#    width = 25, height = 4)
ggplot(ncore_p2,aes(y = n, x = year, group= park, col=park)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()
```


## Interspersion and Juxtaposition index
```{r}
# Interspersion and Juxtaposition index (percent)
#svg(glue("iji.#svg"), 
#    width = 25, height = 4)
ggplot(metrics_tab,aes(y = IJI, x = years, group= park_list, col=park_list)) + 
  facet_grid(~park_list) +
  geom_point(alpha = 0.3) + 
  #geom_hline(yintercept=0, linetype="dashed") +
  theme_bw()+
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()
```


## Contagion index
```{r}
# Contagion index (percent) -------------------------------------------------------------------
#svg(glue("CI.#svg"), 
#    width = 25, height = 4)
ggplot(metrics_tab,aes(y = CI, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()
```


## Total edge
```{r}
# Total edge (m) -------------------------------------------------------------------
#svg(glue("TE.#svg"), 
#    width = 25, height = 4)
ggplot(metrics_tab,aes(y = TE, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()
```


## Mean Core Area Index
```{r}
## cai  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, cai)

#svg(glue("cai.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = cai, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()
```


## Patch Cohesion Index
```{r}
## coh  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, coh)

#svg(glue("coh.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = coh, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

```


## Mean Euclidian Nearest Neighbor Index
```{r}
## enn  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, enn)

#svg(glue("enn.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = enn, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

```


## Aggregation Index
```{r}
## ai  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, ai)

#svg(glue("ai.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = ai, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

```


## Normalized Landscape Shape Index
```{r}
## nlsi  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, nlsi)

#svg(glue("nlsi.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = nlsi, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

```


## Perimeter Area Fractal Dimension
```{r}
## pafrac  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, pafrac)

#svg(glue("pafrac.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = pafrac, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

```


## Clumpy Index
```{r}
## clu  
plot_tab <- metrics_tab %>% 
  dplyr::select(park_list, years, clu)

#svg(glue("clu.#svg"), 
#    width = 25, height = 4)
ggplot(plot_tab,aes(y = clu, x = years, group= park_list, col=park_list)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.major = element_line(colour = "lightgray")) +
  geom_point(alpha = 0.6) + 
  facet_grid(~park_list) +
  geom_hline(yintercept=0, linetype="dashed") +
  theme(text = element_text(size = 25),legend.position = "none")
#dev.off()

```